Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package

### 2020-08-27 paper.tex
update

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package},
            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
            pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother

\title{Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package}
\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
\date{}

@@

@@

We wish to note that challenge facing the analyst using LPA is that these
parameters and the distinct model parameterizations that can be estimated is the
different terminology used. As one example, Scrucca et al. (2017) refer to these
parameterizations not in terms of whether and how parameters are estimated, but
rather in terms of the geometric properties of the distributions that result
from particular parameterizations. Muthen and Muthen (1997-2017) and others
(Pastor et al., 2007) commonly refer to local independence to mean that the
covariances are fixed to zero (also described as the specification of the
covariance matrix as ``diagonal,'' because only the diagonal components, or the
variances, are estimated).

@@

In R, a number of tools \emph{can} be used to carry out LPA. OpenMx can be used for
this purpose (and to specify almost any model possible to specify within a
latent variable modeling approach). However, while OpenMx is very flexible, it
can also be challenging to use. Other tools in R allow for estimating Gaussian
mixture models, or models of multivariate Gaussian (or normal) distributions. In
addition to following the same general approach, using tools that are designed
for Gaussian mixture modeling have other benefits, some efficiency-related (see
RMixMod, which uses compiled C++ code) and others in terms of ease-of-use (i.e.,
the plot methods built-in to RMixMod, mclust, and other tools). However, they
also have some drawbacks, in that it can be difficult to translate between the
model specifications, which are often described in terms of the geometric
properties of the multivariate distributions being estimated (i.e., ``spherical,
equal volume''), rather than in terms of whether and how the means, variances,
and covariances are estimated. They also may use different default settings
(than those encountered in MPlus) in terms of the expectation-maximization
algorithm, which can make comparing results across tools challenging.

@@

One way that tidyLPA is designed to be easy to use is that it assumes a ``tidy''
data structure (Wickham, 2014). This means that it emphasizes the use of a data
frame as both the primary input and output of functions for the package. Because
data is passed to and returned (in amended form, i.e., with the latent profile
probabilities and classes appended to the data) from the function, it makes it
easy to create plots or use results in subsequent analyses. Another noteworthy
feature of tidyLPA is that it provides the same functionality through two
different tools, one that is open-source and available through R, the mclust
package (Scrucca et al., 2017) and one that is available through the commercial
software MPlus (Muthen \& Muthen, 1997-2017). Moreover, as both tools use the
same maximum likelihood estimation procedure, they are benchmarked to produce
the same output. Also, note that we have described the model
specifications with descriptions of what is estimated in terms of the variances
and covariances, the common names for the models (i.e., class-varying
unrestricted), and the covariance matrix associated with the parameterizationfor
the six models that are possible to be estimated on the website for tidyLPA (see
here).

@@

@@

In these examples, we pass the results of one function to the next by \emph{piping}
(using the \texttt{\%\textgreater{}\%} operator, loaded from the \texttt{dplyr} package). We pass the data to
a function that selects relevant variables, and then to \texttt{estimate\_profiles}:

\begin{verbatim}
## tidyLPA analysis using mclust: 
## 
##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
##  1     3       158594.55 158898.38 0.82    0.91     0.92     0.15  0.53  0.01
\end{verbatim}

@@

We can use Mplus simply by changing the package argument for
\texttt{estimate\_profiles()}:

A simple summary of the analysis is printed to the console.
The resulting object can be further passed down a pipeline to
other functions, such as \texttt{compare\_solutions}, \texttt{get\_data}, \texttt{get\_fit},
etc. This is the ``tidy'' part, in that the function can be embedded in a tidy
analysis pipeline.

If you have Mplus installed, you can call the version of this function that uses
MPlus in the same way, by adding the argument \texttt{package\ =\ "MplusAutomation}.

\hypertarget{plotting-the-profiles}{%
\subsubsection{Plotting the profiles}\label{plotting-the-profiles}}

We can plot the profiles by piping the output to \texttt{plot\_profiles()}.

\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}

\hypertarget{model-specification}{%
\subsection{Model specification}\label{model-specification}}

In addition to the number of profiles (specified with the \texttt{n\_profiles}
argument), the model can be specified in terms of whether and how the variable
variances and covariances are estimated.

The models are specified by passing arguments to the \texttt{variance} and \texttt{covariance}
arguments. The possible values for these arguments are:

\begin{itemize}
\tightlist
\item
  \texttt{variances}: ``equal'' and ``zero''
\item
  \texttt{covariances}: ``varying'', ``equal'', and ``zero''
\end{itemize}

@@

We can also run the same process using mplus:

@@

\begin{verbatim}
## # A tibble: 700 x 9
##    model_number classes_number broad_interest enjoyment self_efficacy Class
##           <dbl>          <int>          <dbl>     <dbl>         <dbl> <dbl>
##  1            1              3            3.8       4            1        1
##  2            1              3            3         3            2.75     3
##  3            1              3            1.8       2.8          3.38     3
##  4            1              3            1.4       1            2.75     2
##  5            1              3            1.8       2.2          2        3
##  6            1              3            1.6       1.6          1.88     3
##  7            1              3            3         3.8          2.25     1
##  8            1              3            2.6       2.2          2        3
##  9            1              3            1         2.8          2.62     3
## 10            1              3            2.2       2            1.75     3
## # … with 690 more rows, and 3 more variables: Class_prob <int>,
## #   Probability <dbl>, id <int>
\end{verbatim}

@@

\hypertarget{htmlwidget-a6cf205f9fe062a59d12}{}

@@

First, we note that developing a tool that emphasized both rigor and access has
been, so far, well-received. This suggests that there is an audience (or a
market) for tools to carry out sophisticated tools that may be otherwise
out-of-reach. Tools do not need to be hard to use, a point made by developers
of, for example, tools to carry out highly-sophisticated Bayesian methods using
the most cutting-edge tools (Bürkner \& others, 2017, @makowski2019). We think
that tidyLPA speaks to this, too.

@@

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
            pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\shorttitle{tidyLPA}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{4} University of Oregon\\\textsuperscript{5} }
\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: }
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

\title{Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package}
\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
\date{}

@@

\begin{verbatim}
## Warning: package 'tibble' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'purrr' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## Warning: package 'dplyr' was built under R version 3.6.2
\end{verbatim}

\begin{verbatim}
## NULL
\end{verbatim}

@@

We wish to note that challenge facing the analyst using LPA is that these
parameters and the distinct model parameterizations that can be estimated is the
different terminology used. As one example, Scrucca et al. (2017) refer to these
parameterizations not in terms of whether and how parameters are estimated, but
rather in terms of the geometric properties of the distributions that result
from particular parameterizations. Muthen and Muthen (1997-2017) and others
(Pastor et al., 2007) commonly refer to local independence to mean that the
covariances are fixed to zero (also described as the specification of the
covariance matrix as \enquote{diagonal,} because only the diagonal components, or the
variances, are estimated).

@@

In R, a number of tools \emph{can} be used to carry out LPA. OpenMx can be used for
this purpose (and to specify almost any model possible to specify within a
latent variable modeling approach). However, while OpenMx is very flexible, it
can also be challenging to use. Other tools in R allow for estimating Gaussian
mixture models, or models of multivariate Gaussian (or normal) distributions. In
addition to following the same general approach, using tools that are designed
for Gaussian mixture modeling have other benefits, some efficiency-related (see
RMixMod, which uses compiled C++ code) and others in terms of ease-of-use (i.e.,
the plot methods built-in to RMixMod, mclust, and other tools). However, they
also have some drawbacks, in that it can be difficult to translate between the
model specifications, which are often described in terms of the geometric
properties of the multivariate distributions being estimated (i.e., \enquote{spherical,
equal volume}), rather than in terms of whether and how the means, variances,
and covariances are estimated. They also may use different default settings
(than those encountered in MPlus) in terms of the expectation-maximization
algorithm, which can make comparing results across tools challenging.

@@

One way that tidyLPA is designed to be easy to use is that it assumes a \enquote{tidy}
data structure (Wickham, 2014). This means that it emphasizes the use of a data
frame as both the primary input and output of functions for the package. Because
data is passed to and returned (in amended form, i.e., with the latent profile
probabilities and classes appended to the data) from the function, it makes it
easy to create plots or use results in subsequent analyses. Another noteworthy
feature of tidyLPA is that it provides the same functionality through two
different tools, one that is open-source and available through R, the mclust
package (Scrucca et al., 2017) and one that is available through the commercial
software MPlus (Muthen \& Muthen, 1997-2017). Moreover, as both tools use the
same maximum likelihood estimation procedure, they are benchmarked to produce
the same output. Also, note that we have described the model
specifications with descriptions of what is estimated in terms of the variances
and covariances, the common names for the models (i.e., class-varying
unrestricted), and the covariance matrix associated with the parameterizationfor
the six models that are possible to be estimated on the website for tidyLPA (see
here).

@@

\begin{verbatim}
## Warning: package 'tidyLPA' was built under R version 3.6.2
\end{verbatim}

@@

In these examples, we pass the results of one function to the next by \emph{piping}
(using the \texttt{\%\textgreater{}\%} operator, loaded from the \texttt{dplyr} package). We pass the data to
a function that selects relevant variables, and then to \texttt{estimate\_profiles}.

A simple summary of the analysis is printed to the console.
The resulting object can be further passed down a pipeline to
other functions, such as \texttt{compare\_solutions}, \texttt{get\_data}, and \texttt{get\_fit}.

We can see a number of statistics for each of the models that were estimated, including the
AIC (Aikake information criterion; based on -2 log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size). Both are \emph{penalized} likelihood statistics; the log-likelihood will always decrease with the addition of additional
paremeters estimated by the model, whereas the AIC and BIC will account for the number of parameters; lower values of
either indicate better fit.

Entropy. \texttt{prob\_min} refers to the minimum of the diagonal of the average latent class probabilities for most likely class membership, by assigned class. The minimum should be as high as possible, reflecting greater classification certainty (cases are assigned to classes they have a high probability of belonging to; see Jung \& Wickrama, 2008); \texttt{prob\_max}
refers to the maximum. Finally. \texttt{n\_min} refers to the proportion of the sample assigned to the smallest class (based on most likely class membership); \texttt{n\_max} to the largest. Finally, \texttt{BLRT\_p} refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT
test statistic, can be obtained with the \texttt{get\_fit()} function:

\begin{verbatim}
## # A tibble: 10 x 18
##    Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC
##    <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
##  1     1       1 -81245. 1.63e5 1.63e5 1.63e5 1.63e5 1.62e5 1.63e5 1.63e5
##  2     1       2 -79727. 1.60e5 1.60e5 1.60e5 1.60e5 1.59e5 1.60e5 1.60e5
##  3     1       3 -79239. 1.59e5 1.59e5 1.59e5 1.59e5 1.58e5 1.59e5 1.59e5
##  4     1       4 -79091. 1.58e5 1.59e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
##  5     1       5 -79029. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
##  6     1       6 -78971. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
##  7     1       7 -78930. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
##  8     1       8 -78852. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
##  9     1       9 -78822. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
## 10     1      10 -78751. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
## # ... with 8 more variables: ICL <dbl>, Entropy <dbl>, prob_min <dbl>,
## #   prob_max <dbl>, n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
\end{verbatim}

@@

We can use Mplus simply by changing the package argument for
\texttt{estimate\_profiles()} to \enquote{mplus} or \enquote{MplusAutomation} (either will work; not run here):

These are the very basics; we will return to what we do with this output later in the tutorial.

\hypertarget{specifying-the-model}{%
\subsubsection{\texorpdfstring{Specifying the \emph{model}}{Specifying the model}}\label{specifying-the-model}}

We passed the number of profiles, but no details about the model we wished to estimate.

There are a number of different types of models for LPA that estimate different parameters; choosing
between them is an important consideration, and one that has a bearing on how we will interpret the results we saw above.

Thus, in addition to the number of profiles (specified with the \texttt{n\_profiles}
argument), the model can be specified in terms of whether and how the variable
variances and covariances are estimated. The models are specified by passing arguments to the \texttt{variance} and \texttt{covariance}
arguments. The possible values for these arguments are:

\begin{itemize}
\tightlist
\item
  \texttt{variances}: \enquote{equal} and \enquote{zero}
\item
  \texttt{covariances}: \enquote{varying}, \enquote{equal}, and \enquote{zero}
\end{itemize}

In general, the approach to choosing the model is similar to choosing the number of profiles, requiring \textbf{deciding on the basis of evidence from multiple sources}, including information criteria, statistical tests, and concerns of interpretability and parsimony. The article by \href{https://www.sciencedirect.com/science/article/pii/S0361476X06000543}{Pastor and colleagues (2007)} has helpful information on the model specifications. Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated.

Note that \emph{p} represents different profiles and each parameterization is represented by a 4 x 4 covariance matrix and therefore would represent the parameterization for a four-profile solution. In all of the models, the means are estimated freely in the different profiles. Imagine that each row and column represents a different variable, i.e., the first row (and column) represents broad interest, the second enjoyment, the third self-efficacy, and the fourth another variable, i.e., future goals and plans.

\hypertarget{equal-variances-and-covariances-fixed-to-0-model-1}{%
\subsubsection{1. Equal variances, and covariances fixed to 0 (model 1)}\label{equal-variances-and-covariances-fixed-to-0-model-1}}

In this model, which corresponds to the mclust model wit the name \enquote{EEI}, the variances are estimated to be equal across profiles, indicated by the absence of a p subscript for any of the diagonal elements of the matrix. The covariances are constrained to be zero, as indicated by the 0's between every combination of the variables.

It is specified with \texttt{variances\ =\ "equal"} and \texttt{covariances\ =\ "zero"}.

This model is highly constrained but also parsimonious: the profiles are estimated in such a way that the variables' variances are identical for each of the profiles, and the relationships between the variables are not estimated. In this way, less degrees of freedom are taken used to explain the observations that make up the data. However, estimating more parameters--as in the other models--may better explain the data, justifying the addition in complexity that their addition involves (and their reduction in degrees of freedom). This model is sometimes referred to as a \emph{class-invariant} parameterization.

\[
\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2 }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3 }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] 
\]

\hypertarget{varying-variances-and-covariances-fixed-to-0-model-2}{%
\subsubsection{2. Varying variances and covariances fixed to 0 (model 2)}\label{varying-variances-and-covariances-fixed-to-0-model-2}}

This model corresponds to the mclust model \enquote{VVI} and allows for the variances to be freely estimated across profiles. The covariances are constrained to zero.

It is specified with \texttt{variances\ =\ "varying"} and \texttt{covariances\ =\ "zero"}.

Thus, it is more flexible (and less parsimonious) than model 1, but in terms of the covariances, is more constrained than model 2. This model is sometimes referred to as a \emph{class-varying diagonal} parameterization.

\[ 
\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2p }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3p }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right] 
\]

\hypertarget{equal-variances-and-equal-covariances-model-3}{%
\subsubsection{3. Equal variances and equal covariances (model 3)}\label{equal-variances-and-equal-covariances-model-3}}

This model corresponds to the mclust model \enquote{EEE}. In this model, the variances are still constrained to be the same across the profiles, although now the covariances are estimated (but like the variances, are constrained to be the same across profiles).

It is specified with \texttt{variances\ =\ "equal"} and \texttt{covariances\ =\ "equal"}.

Thus, this model is the first to estimate the covariance (or correlations) of the variables used to create the profiles, thus adding more information that can be used to better understand the characteristics of the profiles (and, potentially, better explain the data). This model is sometimes referred to as a \emph{class-invariant unrestricted} parameterization.

\[
\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] 
\]

\hypertarget{varying-means-varying-variances-and-equal-covariances-model-4}{%
\subsubsection{4. Varying means, varying variances, and equal covariances (model 4)}\label{varying-means-varying-variances-and-equal-covariances-model-4}}

This model, which specifies for the variances to be freely estimated across the profiles and for the covariances to be estimated to be equal across profiles, extends model 3.

It is specified with \texttt{variances\ =\ "varying"} and \texttt{covariances\ =\ "equal"}.

Unfortunately, this model cannot be specified with mclust, though it can be with MPlus; this model \emph{can} be used with the functions to interface to MPlus described below.

\[
\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right] 
\]

\hypertarget{varying-means-equal-variances-and-varying-covariances-model-5}{%
\subsubsection{5. Varying means, equal variances, and varying covariances (model 5)}\label{varying-means-equal-variances-and-varying-covariances-model-5}}

This model specifies the variances to be equal across the profiles, but allows the covariances to be freely estimated across the profiles.

It is specified with \texttt{variances\ =\ "equal"} and \texttt{covariances\ =\ "varying"}.

Like model 4, this model cannot be specified with mclust, though it can be with MPlus. Again, this model \emph{can} be used with the functions to interface to MPlus described below.

\[
\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] \quad 
\]

\hypertarget{varying-variances-and-varying-covariances-model-6}{%
\subsubsection{6. Varying variances and varying covariances (model 6)}\label{varying-variances-and-varying-covariances-model-6}}

This model corresponds to the mclust model \enquote{VVV}. It allows the variances and the covariances to be freely estimated across profiles.

It is specified with \texttt{variances\ =\ "varying"} and \texttt{covariances\ =\ "varying"}.

Thus, it is the most complex model, with the potential to allow for understanding many aspects of the variables that are used to estimate the profiles and how they are related. However, it is less parsimonious than all of the other models, and the added parameters should be considered in light of how preferred this model is relative to those with more simple specifications. This model is sometimes referred to as a \emph{class-varying unrestricted} parameterization.

\[
\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right] 
\]

@@

\begin{verbatim}
## Warning: 
## One or more analyses resulted in warnings! Examine these analyses carefully: model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
\end{verbatim}

\begin{verbatim}
## Warning: 
## One or more analyses resulted in warnings! Examine these analyses carefully: model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
\end{verbatim}

The results of this analysis suggests that the Model 6, two or seven class models are best.

@@

@@

\hypertarget{htmlwidget-31a3e9100620ea87273d}{}

@@

First, we note that developing a tool that emphasized both rigor and access has
been, so far, well-received. This suggests that there is an audience (or a
market) for tools to carry out sophisticated tools that may be otherwise
out-of-reach. Tools do not need to be hard to use, a point made by developers
of, for example, tools to carry out highly-sophisticated Bayesian methods using
the most cutting-edge tools (Bürkner \& others, 2017, Makowski, Ben-Shachar, Chen, and Lüdecke (2019)). We think
that tidyLPA speaks to this, too.

@@

\leavevmode\hypertarget{ref-burkner2017}{}%
Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
